Model Checking

### 2023-07-21 owen/Write Ups/03_existing_models.qmd
quick update to instead of fitting to every game, construct the windows, only take the final game in each window, use that rather than all games. plotted the parameter estimates and did posterior predictive checking. seems to work reasonably well still

knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 8)

@@

alpha_mean <- apply(alpha_draws, 2, mean)
beta_mean <- apply(beta_draws, 2, mean)
gamma_mean <- apply(gamma_draws, 2, mean)
## then get the id's from this, based on the names
names(alpha_mean)

@@

>>>>

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8)

@@

alpha_mean <- apply(alpha_draws, 2, mean)
beta_mean <- apply(beta_draws, 2, mean)
gamma_mean <- apply(gamma_draws, 2, mean)
## then get the id's from this, based on the names
# names(alpha_mean)

@@

## Updates July

Only consider a single game in each window?

get_hist_mod <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)),
           window_id = row_number()/prev_n)

  hist_games %>% filter(window_id - round(window_id) == 0)
}

tidy_games_mod <- map_dfr(top_players, get_hist_mod, bullet_60, prev_n = 10) %>% 
  as_tibble()

## so this takes windows of size 10, so keep 10th game, 
## use history from the previous 9 games each time

tidy_games_mod

init_data_mod <- tidy_games_mod %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
  filter(focal_result != 0.5)

Then quickly fit model to this data.

```{r}
stan_data_mod <- list(N = nrow(init_data_mod),
                  J = length(top_players),
                  y = init_data_mod$focal_result,
                  id = init_data_mod$focal_id,
                  colour = init_data_mod$focal_white,
                  elo = init_data_mod$elo_diff,
                  win_prop = init_data_mod$ave_prop)

fit3_mod <- mod$sample(data = stan_data_mod,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)
```

```{r}
fit3_mod$summary()

mcmc_hist(fit3_mod$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
```

Get the posterior predictive checks.

```{r}
fit_samples <- as_draws_df(fit3_mod$draws())

fit_samp <- fit_samples %>% 
  select(!starts_with(c("log_lik", "yrep")))

y_rep <- fit_samples %>% select(starts_with("y_rep"))

rm(fit_samples)
rm(fit_samp)

### then want to join this with the correct input, i.e player, etc

y_rep_mod <- y_rep %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from=rowname, values_from=value) %>% 
  ## remove the draws and chains here
  mutate(focal_id = stan_data_mod$id)

games_won <- y_rep_mod %>% 
  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
  group_by(focal_id, draw) %>% 
  summarise(games_won = sum(y))

orig_data <- tibble(outcome = stan_data_mod$y,
                    focal_id = stan_data_mod$id)

orig_games_won <- orig_data %>% 
  group_by(focal_id) %>% 
  summarise(games_won = sum(outcome))

games_won %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram(bins = 20) +
  facet_wrap(~focal_id, scales = "free") +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  labs(title = "Final Outcome in Window Only")
